---
title: "EX. RESEARCH DESIGN FINAL PROJECT"
author: "Della Gatta Antonio"
date: ""
output: 
  html_document: 
    code_folding: hide
    theme: journal
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
#install.packages("palmerpenguins")
library(palmerpenguins)
library(ggplot2)
library(ggpubr)
library(dplyr)
```

# Palmer Penguin data

### dataset

The Palmer penguins dataset by Allison Horst, Alison Hill, and Kristen Gorman was first made publicly available as an R package.

Size measurements, clutch observations, and blood isotope ratios for 344 adult foraging Adelie, Chinstrap, and Gentoo penguins observed on islands in the Palmer Archipelago near Palmer Station, Antarctica.

The variables in the dataset are:

-   **species**

-   **island**

-   **bill length**

-   **bill depth**

-   **flipper length**

-   **body max**

-   **sex**

-   **year**

```{r, message=FALSE, warning=FALSE,  comment=""}

dat<- penguins
dat <- na.omit(dat)

dat<- dat[-c(259,260,263), ]

#View(penguins)
#View(dat)

dat
```

### bill length by sex

As a first analysis I would focus my attention on the male and female penguins of the "Adelie" species

Let's test with a two sample unpaired t-test

H0 : there's no difference between the bill length of the male and female penguins *(μM=μF)*

H1: there's difference between the bill length of the male and female penguins *(μM≠μF)*

**display our data\
**

```{r, warning=FALSE, message=FALSE,  comment=""}
library(tseries)
library(dplyr)

new_male <- dat %>%
  filter(species == 'Adelie' & sex == 'male')

new_female <- dat %>%
  filter(species == 'Adelie' & sex == 'female')


male<- new_male[,3] #select only value of interest
male<- as.numeric(unlist(male)) #eliminate list

female<- new_female[,3] #select only value of interest
female<- as.numeric(unlist(female)) #eliminate list
```

Let's display our data in order to see graphically if there are already some difference.

```{r, warning=FALSE, message=FALSE}

par(mfrow=c(1,2))
boxplot(male, main ="male", col ="orange")
boxplot(female, main = "female", col ="yellow")
```

**check for normality**

```{r, warning=FALSE, message=FALSE, comment=""}

shapiro.test(male)



shapiro.test(female)

```

since that the value is not less than 0.5 we can say that, our data, are normally distributed.

**we can also display it graphically**

```{r, message=FALSE, warning=FALSE, comment=""}

par(mfrow=c(1,2))

x1<- hist(male, col='steelblue')
x2<- hist(female,col='steelblue')



```

**check for "equal" variances**

sd(male)/sd(female) should not be greater than 2.

In this case we'll use an F- test.\\

```{r, warning=FALSE, message=FALSE,  comment=""}


var.test(male, female, 
         alternative = "two.sided")
```

The p-value of **F-test** is p = 0.3296 which is greater than the significance level 0.05. In conclusion, there is no significant difference between the two variances.

So we can say that there's homogeneity of variances

### check for outliers

```{r, message=FALSE, warning=FALSE,  comment=""}
par(mfrow=c(1,2))
boxplot(male,main = "male", col ="orange")
outliers_values=boxplot.stats(male)
outliers_values

boxplot(female,main = "female", col ="yellow")
outliers_values=boxplot.stats(female)$out
outliers_values
```

### remove outliers for male

```{r, message=FALSE, warning=FALSE,  comment=""}
remove_outliers_iqr=function(x){
  IQR=summary(x)[5]-summary(x)[2]
  up_out=summary(x)[5]+1.5*IQR
  down_out=summary(x)[2]-1.5*IQR
  out_data=c(which(x<down_out), which(x>up_out))
  x_new=x[-out_data]
  print(length(x_new))
  boxplot(x_new)
  summary(x_new)}



remove_outliers_iqr(male)

```

now let's perform our test

```{r, warning=FALSE, message=FALSE,  comment=""}
source('http://www.sthda.com/upload/rquery_t_test.r')
rquery.t.test(male, female, paired=FALSE)
```

There aren't enough evidence to claim that the null hypothesis is true.

The means of the bill length of the penguins are different

## Difference of flipper length by species and sex

Now we can focus our research on the difference of flipper length of three species of penguins.

We can also divide them by sex. The test that we can use is a two way anova.

### Two-way ANOVA test hypotheses

1.  There is no difference in the means of first factor

2.  There is no difference in means of second factor

3.  There is no interaction between factors 

The alternative hypothesis for cases 1 and 2 is: the means are not equal.

The alternative hypothesis for case 3 is: there is an interaction between factor

let's see our data

```{r, message=FALSE, warning=FALSE, comment=""}
str(dat)
```

**Balanced design?**

```{r, message=FALSE, warning=FALSE,  comment=""}
table(dat$species, dat$sex)

```

**display our data**

```{r, message=FALSE, warning=FALSE,  comment=""}

library("ggpubr")

ggboxplot(dat, x = "species", y = "flipper_length_mm", color = "sex",
          palette = c("#00AFBB", "#E7B800"))

ggline(dat, x = "species", y = "flipper_length_mm", color = "sex",
       add = c("mean_se", "dotplot"),
       palette = c("#00AFBB", "#E7B800"))


```

Remove outliers

there aren't significant outliers

```{r, message=FALSE, warning=FALSE, comment=""}
#find Q1, Q3, and interquartile range for values in column A

df<- dat

Q1 <- quantile(df$flipper_length_mm, .25)
Q3 <- quantile(df$flipper_length_mm, .65)
IQR <- IQR(df$flipper_length_mm)

#only keep rows in dataframe that have values within 1.5*IQR of Q1 and Q3
no_outlier <- subset(df, df$flipper_length_mm > (Q1 - 1.5*IQR) & df$flipper_length_mm < (Q3 + 1.5*IQR))

#view row and column count of new data frame
print("dimension dataset always equal : 330   8")
#dim(no_outlier) 



```

**Check for normality**

```{r, message=FALSE, warning=FALSE, comment=""}
library(rstatix)

 no_outlier %>%
  group_by(species,sex) %>%
  shapiro_test(flipper_length_mm)

```

the p-value is never lower than 0.05

-   normality checked

------------------------------------------------------------------------

**now let's run our test**

Two-way interaction plot

```{r, warning=FALSE, message=FALSE,  comment=""}

library(stats)
interaction.plot(x.factor = no_outlier$species, trace.factor = no_outlier$sex, 
                 response = no_outlier$flipper_length_mm, fun = mean, 
                 type = "b", legend = TRUE, 
                 xlab = "species", ylab="flipper_length_mm",
                 pch=c(1,19), col = c("#00AFBB", "#E7B800"))



```

Two-way ANOVA with interaction effect

```{r, message=FALSE, warning=FALSE, comment=""}
mod2 <- aov(flipper_length_mm ~ sex + species, data = no_outlier)
summary(mod2)

```

Two-way ANOVA without interaction effect

```{r, warning=FALSE, message=FALSE,  comment=""}


mod3 <- aov(flipper_length_mm ~ sex * species, data = no_outlier)
summary(mod3)

```

**check for normality of the residuals**

first plot

```{r, warning=FALSE, message=FALSE,  comment=""}
# 2. Normality
plot(mod3, 2)
```

then use the shapiro test

```{r, warning=FALSE, message=FALSE,  comment=""}
# Extract the residuals
aov_residuals <- residuals(object = mod3)
# Run Shapiro-Wilk test
shapiro.test(x = aov_residuals )
```

-   checked

**check for HOMOSCEDASTICITY**

first plot residuals

```{r, warning=FALSE, message=FALSE,echo=FALSE}


plot(mod3, 1)
```

Use the **Levene's test** to check the **homogeneity of variances**. The function **leveneTest**() [in **car** package] will be used:

```{r, warning=FALSE, message=FALSE}
library(car)
library(PMCMRplus)
leveneTest(flipper_length_mm ~ species * sex, data = no_outlier)



```

try another test

```{r, warning=FALSE, message=FALSE}
hartleyTest(flipper_length_mm ~ interaction(sex,species), data = no_outlier)


```

```{r, warning=FALSE, message=FALSE}
library(onewaytests)

welch.test(flipper_length_mm ~ interaction(sex,species), data = no_outlier)
```

-   **homogeneity of variances not present**

## bill length by island

Another interesting aspect was to understand if the bill length could have changed from an island to another.

So I've completed this task thanks to a non parametric test

```{r, warning=FALSE, message=FALSE,  comment=""}
#View(dat)
new_m <- dat %>%
  filter(island == 'Dream')


new_f <- dat %>%
  filter(island == 'Biscoe')


dream<- new_m[1:100,3] #select only value of interest
dream<- as.numeric(unlist(dream)) #eliminate list

biscoe<- new_f[1:100,3] #select only value of interest
biscoe<- as.numeric(unlist(biscoe)) #eliminate list
```

check normality

```{r, comment="", warning=FALSE, message=FALSE}
shapiro.test(dream)

shapiro.test(biscoe)
```

the p value is lower than 0.05 so, in this case, the value of the bill length is not normally distributed.

-   check variances

```{r, warning=FALSE, message=FALSE,, comment=""}
fligner.test(dream ~ biscoe)
```

### does the median value of the penguins on the first island differ from the median of the penguin on the second one?

**display our data**

```{r, warning=FALSE, message=FALSE,  comment=""}
library(tidyverse)
library(rstatix)
library(ggpubr)

my_data <- data.frame( 
  group = rep(c("dream", "biscoe"), each = 100),
  weight = c(dream,  biscoe)
)


ggboxplot(my_data, x = "group", y = "weight", 

          color = "group", palette = c("red", "pink"),

          ylab = "Weight", xlab = "Groups")


```

**run our test**

```{r, message=FALSE, warning=FALSE, , comment=""}
res <- wilcox.test(dream , biscoe)

res
```

it doesn't seem that there's a real difference.

**effect size**

```{r, warning=FALSE, message=FALSE,  comment=""}
library(coin)
my_data %>% wilcox_effsize(weight ~ group)
```

The Wilcoxon test showed that the difference was not significant (p \> 0.05, effect size r = 0.0313).


